Learning Discrete-valued Bayesian Networks from Mixed Data

### 2015-11-18 BN_discretizer_paper.tex
Update experiment section

\begin{table}[h]
  \centering
  \input{table_autoMPG_1.tex}
  \caption{Discretization result of Auto MPG dataset based on the graph in Fig. \ref{fig:auto_graph_1}. The discretization policy for the heuristic matches that for the optimal Bayesian approach. The MDL method is less sensitive to distribution of values of other variables, thus it fails to return discretization edges.
  \label{table:auto_disc_table_1}
\end{table}

@@

 %{\centering
%\begin{minipage}{0.50\textwidth}
%  \centering
%  \scalebox{0.7}{
%  \input{graph_autocar_2}}
%  \captionof{figure}{The learned discrete-\newline valued Bayesian network}
%  \label{fig:auto_graph_2}
%\end{minipage}
%\begin{minipage}{0.50\textwidth}
%  \raggedleft
%  \input{table_autoMPG_2.tex}
%  \captionof{table}{The discretization edges of variables in Fig \ref{Auto_graph_2}}
%  \label{Auto_MPG_disc_table_2}
%\end{minipage}
%}

@@

The Bayesian network in Figure \ref{fig:wine_graph_1} is obtained by equal-width discretizing each continuous variable into 3 intervals and running K2 a thousand times. Table \ref{table:wine_disc_table_1} shows a comparison between the optimal Bayesian approach and the MDL principle discretization method.
The discretization result obtained by the optimal Bayesian approach is better than the MDL discretization method in terms of likelihood, although the latter returns non-trivial result in this experiment. Note that the 
heuristic method discretizes variable $G$, $H$ and $J$ into one more interval than the optimal Bayesian approach, and has less number of intervals on variable $M$. Although the heuristic method has more intervals after discretization on all variables, it does not lead to a higher likelihood result. This can be interpreted as that the extra discretization edges on $G$, $H$ and $J$ are less necessary and the losing discretization edge on $M$ is important.
In addition, as shown in next dataset, the heuristic method might cut continuous variables into too many intervals. Therefore, the optimal Bayesian approach should be considered a better method if the running time is not a concern.

@@

\begin{table}
  \centering
  \scalebox{1.0}{
  \input{table_wine_1.tex}
  }
  \caption{Discretization result of Wine dataset based on the graph Fig. \ref{fig:wine_graph_1}
  \todo{MDL has bins, heuristic is slightly worse in logl but still good. MDL keeps all of its CPDs small, no neighboring vars both have more than 1 interval.}
  \tim{Should we scale the logl? What is the best way to do that?}
  }
  \label{table:wine_disc_table_1}
\end{table}

@@

\begin{figure}[ht]
  \centering
    \input{graph_wine_exp1_distr}
  \caption{Comparison of the discretization results on variable $A$ and $C$ obtained by the proposed method and the MDL method}
  \label{fig:wine_exp1_distr}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \begin{tabular}{cc}
  \centering
  \scalebox{0.7}{
    \input{graph_wine_2}}
  \end{tabular}
  \caption{The learned discrete-valued Bayesian network from the Wine dataset \todo{using optimal Bayesian method}}
  \label{fig:wine_graph_2}
\end{figure}

@@

The Bayesian network structure in Figure \ref{fig:housing_graph_1} is obtained by prediscretizing each continuous variable into five uniform-width intervals, then running the K2 algorithm one thousands time and choosing the network with the highest K2 score. Table \ref{table:housing_disc_table_1} shows the numbers of intervals after discretization on each continuous variables. MDL method does not discretize most variables. The proposed method with the approximation discretizes variables into too many intervals. For example, it discretized $J$ into 58 intervals. The proposed method without the approximation has a good control of number of intervals.
\begin{figure}[ht]
  \centering
  % \begin{tabular}{cc}
    \input{graph_housing_1}
  % \end{tabular}
  \caption{\todo{Structure from prediscretized K2 structure learning to compare the methods for Housing dataset}}
  \label{fig:housing_graph_1}
\end{figure}

@@

\begin{figure}[ht]
    % \begin{tabular}{cc}
      \input{graph_housing_exp1_distr_8_5}
    % \end{tabular}
  \caption{The discretization result on variable $H$ and $E$ of the network in Figure~\ref{fig:housing_graph_1}}
  \label{fig:housing_exp1_distr_8_5}
\end{figure}

@@

\begin{figure}[ht]
    % \begin{tabular}{cc}
      \input{graph_housing_exp2_distr_3_5}
    % \end{tabular}
  \caption{The discretization result on variable $C$ and $E$ of the network in Figure~\ref{fig:housing_graph_2}}
  \label{fig:housing_exp2_distr_3_5}
\end{figure}

>>>>

\begin{table}[h]
  \centering
  \caption{Discretization result of Auto MPG dataset based on the graph in Fig. \ref{fig:auto_graph_1}. The discretization policy for the heuristic matches that for the optimal Bayesian approach. The MDL method is less sensitive to distribution of values of other variables, thus it fails to return discretization edges.}
  \input{table_autoMPG_1.tex}
  \label{table:auto_disc_table_1}
\end{table}

@@

 %{\centering
%\begin{minipage}{0.50\textwidth}
  %\centering
  %\scalebox{0.7}{
  %\input{graph_autocar_2}}
  %\captionof{figure}{The learned discrete-\newline valued Bayesian network}
  %\label{fig:auto_graph_2}
%\end{minipage}
%\begin{minipage}{0.50\textwidth}
  %\raggedleft
  %\input{table_autoMPG_2.tex}
  %\captionof{table}{The discretization edges of variables in Fig \ref{Auto_graph_2}}
 % \label{Auto_MPG_disc_table_2}
%\end{minipage}
%}

@@

The Bayesian network in Figure \ref{fig:wine_graph_1} is obtained by equal-width discretizing each continuous variable into 3 intervals and running K2 a thousand times. Table \ref{table:wine_disc_table_1} shows a comparison between the optimal Bayesian approach and the MDL principle discretization method.
The discretization result obtained by the optimal Bayesian approach is better than the MDL discretization method in terms of likelihood, although the latter returns non-trivial result in this experiment. Note that the 
heuristic method discretizes variable $G$, $H$ and $J$ into one more interval than the optimal Bayesian approach, and has less number of intervals on variable $M$. Although the heuristic method has more intervals after discretization on all variables, it does not lead to a higher likelihood result. This can be interpreted as that the extra discretization edges on $G$, $H$ and $J$ are less necessary and the losing discretization edge on $M$ is important. On the other hand, if running time is a concern, then the heuristic method can be considered a better approach in this case, since its discretization result is nearly as good as the result of the optimal Bayesian approach.

Notice that there are some discretization edges shown in all three discretization methods, such as \num{1.42} and \num{2.35} on variable $C$ and \num{0.785} on variable $L$. This indicates that MDL indeed can find some important discretization edges, but it is not sensitive enough to find other edges.

@@

\begin{table}
  \centering
  \scalebox{1.0}{
  \input{table_wine_1.tex}
  }
  \caption{Discretization result of Wine dataset based on the graph Fig. \ref{fig:wine_graph_1}. The optimal Bayesian approach has the best discretization result in terms of likelihood. The heuristic method obtains a result nearly as good as the Bayesian approach. The MDL method discretizes more than half of continuous variables but still does not discretize enough intervals. Some discretization edges are obtained by all three methods.
  \todo{MDL has bins, heuristic is slightly worse in logl but still good. MDL keeps all of its CPDs small, no neighboring vars both have more than 1 interval.}
  \tim{Should we scale the logl? What is the best way to do that?}
  }
  \label{table:wine_disc_table_1}
\end{table}

@@

%\begin{figure}[ht]
%  \centering
%    \input{graph_wine_exp1_distr}
%  \caption{Comparison of the discretization results on variable $A$ and $C$ obtained by the proposed method and the MDL method}
%  \label{fig:wine_exp1_distr}
%\end{figure}

@@

\begin{figure}[ht]
  \centering
  \scalebox{0.7}{
    \input{graph_wine_2}}
  \caption{The learned discrete-valued Bayesian network from the Wine dataset \todo{using optimal Bayesian method}}
  \label{fig:wine_graph_2}
\end{figure}

@@

The Bayesian network structure in Figure \ref{fig:housing_graph_1} is obtained by prediscretizing each continuous variable into five uniform-width intervals, then running the K2 algorithm one thousands time and choosing the network with the highest K2 score. Table \ref{table:housing_disc_table_1} shows the numbers of intervals after discretization on each continuous variables. MDL method does not discretize most variables. The proposed method with the approximation discretizes variables into too many intervals. For example, it discretized $J$ into 58 intervals. The proposed method without the approximation has a good control of number of intervals.
\begin{figure}[ht]
  \centering
  \begin{tabular}{cc}
   \input{graph_housing_1}
   \end{tabular}
  \caption{\todo{Structure from prediscretized K2 structure learning to compare the methods for Housing dataset}}
  \label{fig:housing_graph_1}
\end{figure}

@@

\begin{figure}[ht]
     %\begin{tabular}{cc}
    \input{graph_housing_exp1_distr_8_5}
     %\end{tabular}
  \caption{The discretization result on variable $H$ and $E$ of the network in Figure~\ref{fig:housing_graph_1}}
  \label{fig:housing_exp1_distr_8_5}
\end{figure}

@@

\begin{figure}[ht]
     %\begin{tabular}{cc}
      \input{graph_housing_exp2_distr_3_5}
     %\end{tabular}
  \caption{The discretization result on variable $C$ and $E$ of the network in Figure~\ref{fig:housing_graph_2}}
  \label{fig:housing_exp2_distr_3_5}
\end{figure}
